IFA Comparison of Noncognitive Skills \\ BCS and MCS

### 2018-05-22 codes/R/paper/cohorts_nc.Rnw
delete incineq, reorder SCL in plot
paper update DO NOT RUN, to be checked

\documentclass[pdftex,10pt,twoside,a4paper]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{amstext}% for resizing the symbol in math
\usepackage{amsthm}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[babel]{csquotes}
\usepackage{booktabs}
\usepackage{tocloft}
\usepackage{dsfont}
\usepackage{bbm}
\usepackage{xcolor}
\usepackage{accents}
\usepackage{setspace} % for spacing
\onehalfspacing
\usepackage{rotating}
\usepackage{fullpage}
\usepackage{pdflscape}  % for landscape page
\usepackage{lmodern}  %font
\usepackage{natbib}
\usepackage{array}
\usepackage{graphicx}
\usepackage{adjustbox}
\usepackage{tikz}
\usepackage{bm}
\usepackage{tabularx}
\usepackage{marvosym}
\usepackage{calc}
\usepackage{array}
\usepackage{verbatim}
\usepackage{multirow}
\usepackage[multiple]{footmisc}   % for multiple footnotes
\usepackage{textgreek}            % for greek letters in text

@@

read_chunk(here("codes/R/1_load.R"))
read_chunk(here("codes/R/2_efa.R"))
read_chunk(here("codes/R/3_inv_main.R"))
read_chunk(here("codes/R/3_inv_aux.R"))
read_chunk(here("codes/R/3_inv_fit.R"))
read_chunk(here("codes/R/4_scores.R"))
read_chunk(here("codes/R/4_scores_aux.R"))
read_chunk(here("codes/R/5_graphs.R"))
read_chunk(here("codes/R/6_decvars.R"))
read_chunk(here("codes/R/7_counterf.R"))
read_chunk(here("codes/R/7_counterf_graphs.R"))
read_chunk(here("codes/R/7_counterf_tabs.R"))
read_chunk(here("codes/R/7_counterf_seqdec.R"))
read_chunk(here("codes/R/getpars.R"))

<<PREAMBLE, cache=FALSE >>=
@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\clearpage

@@

@@

@@

\clearpage
\begin{figure}
<<FACDENS, dev='tikz', fig.height=10, fig.width=10, out.width='1\\linewidth'>>=
@
\caption{Density of Scored Factors}
{\scriptsize \emph{Notes}: The figure shows estimated kernel densities of the scored factors separately by cohort.  \par}
\end{figure}

\clearpage
\begin{figure}
<<RAWHIST, dev='tikz', fig.height=10, fig.width=10, out.width='1\\linewidth'>>=
@
\caption{Frequency of Raw Scores}
{\scriptsize \emph{Notes}: The figure shows frequencies of raw (sum) scores separately by cohort. Raw scores are obtained using the same 11 items in the CFA model, building a categorical index. Children for which each behaviour ``certainly applies/true" are assigned zero points, one point for ``somewhat applies/true" and two points for ``does not apply/not true". For binary items, one point is assigned for ``true". The raw scores are the sum of the points across the 11 behaviours considered, six for EXT and five for INT. \par}
\end{figure}

\textbf{REVIEW} We can derive scores for $\theta^{EXT}$ and $\theta^{INT}$ directly, and plot them against socioeconomic measures observed in the data. Our main socioeconomic indicator is a measure of weekly total family net income in the prices of the interview year.\footnote{This measure is harmonised across cohorts and derived as part of CLOSER, which is funded by the ESRC.} Using the estimated parameters for the model with invariant thresholds and loadings (but non-invariant intercepts), I compute the scores for each individual in the sample (using the Empirical Bayes modal approach). In \autoref{fig:meanci}, we can see that the gradient in noncognitive skills along family income quintiles is more pronounced in the MCS cohort.

<<INEQ_STATS, dependson="FA_SCORES_MAIN", results="markup">>=

# income 
tb <- data.frame(
  aggregate(scores2plot$faminc10_infl, by=list(scores2plot$cohort), FUN = function(x) quantile(x, c(.1, .25, .75, .9), na.rm=T))
)
tb$d9010 <- tb$x[,4]/tb$x[,1]
tb$d7525 <- tb$x[,3]/tb$x[,2]
cat("Quantile differences in income by cohort")
tb

tb.ext <- data.frame(
  aggregate(scores2plot$EXT, by=list(scores2plot$cohortsex), FUN = function(x) quantile(x, c(.1, .25, .5, .75, .9), na.rm=T))
)
tb.ext$d9010 <- abs(tb.ext$x[,"90%"]-tb.ext$x[,"10%"])
tb.ext$d9010 <- abs(tb.ext$x[,"75%"]-tb.ext$x[,"25%"])
tb.ext$d5010 <- abs(tb.ext$x[,"50%"]-tb.ext$x[,"10%"])
tb.ext$d9050 <- abs(tb.ext$x[,"90%"]-tb.ext$x[,"50%"])
cat("\n Quantile differences in EXT by cohort/gender")
tb.ext

tb.int <- data.frame(
  aggregate(scores2plot$INT, by=list(scores2plot$cohortsex), FUN = function(x) quantile(x, c(.1, .25, .5, .75, .9), na.rm=T))
)
tb.int$d9010 <- abs(tb.int$x[,"90%"]-tb.int$x[,"10%"])
tb.int$d9010 <- abs(tb.int$x[,"75%"]-tb.int$x[,"25%"])
tb.int$d5010 <- abs(tb.int$x[,"50%"]-tb.int$x[,"10%"])
tb.int$d9050 <- abs(tb.int$x[,"90%"]-tb.int$x[,"50%"])
cat("\n Quantile differences in INT by cohort/gender")
tb.int

\clearpage
\begin{figure}
\begin{center}
<<INCINEQ, dependson="FA_SCORES_MAIN", dev='tikz', fig.height=4, fig.width=4, out.width='.5\\linewidth'>>=
@
\end{center}
\caption{Evolution of income inequality }
{\scriptsize \emph{Notes}:  \par}
\end{figure}

<<FACINEQ_DEMEAN>>=
# function for demeaning
@

% 
% \begin{figure}
% <<FACINEQ, dependson="FA_SCORES_MAIN", dev='tikz', fig.height=12, fig.width=10, out.width='1\\linewidth'>>=
% @
% \caption{Mean scores across income quintiles}\label{fig:meanci}
% {\scriptsize \emph{Notes}: The figure plots the mean and its 95\% confidence interval across the quintiles in the income distribution. \par}
% \end{figure}

\begin{figure}
<<FACINEQ_SC, dependson="FA_SCORES_MAIN", dev='tikz', fig.height=12, fig.width=10, out.width='1\\linewidth'>>=
@
\caption{Mean scores across social class}\label{fig:meanci_sc}
{\scriptsize \emph{Notes}: The figure plots the mean and its 95\% confidence interval across harmonised social class of the cohort member's family at 10 years of age. \par}
\end{figure}

\begin{figure}
<<FACINEQ_YS, dependson="FA_SCORES_MAIN", dev='tikz', fig.height=12, fig.width=10, out.width='1\\linewidth'>>=
@
\caption{Mean scores across maternal schooling}\label{fig:meanci_ys}
{\scriptsize \emph{Notes}: The figure plots the mean and its 95\% confidence interval across grouped years of schooling of the cohort member's mother at 5 years of age. \par}
\end{figure}

\begin{figure}
<<FACINEQ_ME, dependson="FA_SCORES_MAIN", dev='tikz', fig.height=12, fig.width=10, out.width='1\\linewidth'>>=
@
\caption{Mean scores across maternal employment}\label{fig:meanci_me}
{\scriptsize \emph{Notes}: The figure plots the mean and its 95\% confidence interval by employment status of the cohort member's mother at 5 years of age. \par}
\end{figure}

@@

<<FORMULAS>>=
# define variables and formulas for regression and decomposition
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% 
% \begin{table}
% \caption{Quantifying the gradient}
% \centering
% \resizebox{\textwidth}{!}{
% <<FA_REGS_REAL_M, results='asis'>>=
% # regress scores on income
% library(stargazer)
% 
% decvars_reg <- c(decvars, "cohort") # add cohort
% interact_reg <- paste0(decvars_reg[! decvars_reg %in% c("cohort")], "*cohort")
% 
% form_ext_reg <- as.formula(paste("EXT ~ ", paste0(decvars2, collapse = '+')))
% form.ext2i <- as.formula(paste("EXT ~ ", paste0(c(decvars2, interact2), collapse = '+')))
% form.int2 <- as.formula(paste("INT ~ ", paste0(decvars2, collapse = '+')))
% form.int2i <- as.formula(paste("INT ~ ", paste0(c(decvars2, interact2), collapse = '+')))
% 
% r.ext.m  <- lm(form.ext2,  data=subset(scores2plot, sex=="M"))
% r.exti.m <- lm(form.ext2i, data=subset(scores2plot, sex=="M"))
% r.ext.f  <- lm(form.ext2,  data=subset(scores2plot, sex=="F"))
% r.exti.f <- lm(form.ext2i, data=subset(scores2plot, sex=="F"))
% r.int.m  <- lm(form.int2,  data=subset(scores2plot, sex=="M"))
% r.inti.m <- lm(form.int2i, data=subset(scores2plot, sex=="M"))
% r.int.f  <- lm(form.int2,  data=subset(scores2plot, sex=="F"))
% r.inti.f <- lm(form.int2i, data=subset(scores2plot, sex=="F"))
% 
% stargazer(r.ext.m, r.exti.m, r.ext.f, r.exti.f, r.int.m, r.inti.m, r.int.f, r.inti.f,
%           column.labels=c("M","M","F","F","M","M","F","F"), #,"BCS F","MCS F","BCS F","MCS F"),
%           dep.var.caption = "", multicolumn = FALSE,
%           #omit=c("Constant", paste0("region", c(2:9))),
%           no.space=TRUE,
%           omit.stat= c("rsq", "f", "ser", "aic", "ll"),
%           notes = "All estimates control for Government Office Region of birth.",
%           single.row=TRUE,
%           float=FALSE    # needs begin table outside it
% )
% @
% }
% \end{table}

\clearpage
\subsection{Decomposition of inequality}

We have populations $k \in \mathcal{K}$. For each population, we observe covariates $X_k$. For some $j \subseteq \mathcal{K}$, we observe outcome $Y_j$. We can identify the covariate distributions $F_{X_k}$ and the conditional distributions $F_{Y_j \mid X_j}$ and quantile functions $Q_{Y_j \mid X_j}$. Define the support of $X_k$ as $\mathcal{X}_k$ and the region of interest for the outcome as $\mathcal{Y}_j$.

The interest is in the counterfactual distribution of the outcome:
$$ F_{Y(j \mid k)}(y) = \int_{\mathcal{X}_k} F_{Y_j \mid X_j}(y \mid x) d F_{X_k}(x), \qquad y \in \mathcal{Y}_j $$
and the corresponding quantiles:
$$ Q_{Y(j \mid k)}(\tau) = F^\leftarrow_{Y(j \mid k)}(\tau) $$

Given this, we can define counterfactual effects (CE):
\bi
\item Type 1 (Structure effect): CE of changing the conditional distribution: $Q_{Y(j \mid k)}(\tau) - Q_{Y(l \mid k)}(\tau)$
\item Type 2 (Composition effect): CE of changing the covariate distribution: $Q_{Y(j \mid k)}(\tau) - Q_{Y(j \mid m)}(\tau)$
\item Type 3: CE of changing both conditional and covariate distribution: $Q_{Y(j \mid k)}(\tau) - Q_{Y(l \mid m)}(\tau)$
\ei

In our application, we have $k=BCS,MCS$. We can thus decompose the observed differences in the distribution of non-cognitive skills into structure and composition effects. For example, the difference in the distribution of externalising skills can be written as:
$$F_{(\theta^{EXT}_{MCS} \mid X_{MCS})} - F_{(\theta^{EXT}_{BCS} \mid X_{BCS})} =
  \left[ F_{(\theta^{EXT}_{MCS} \mid X_{MCS})} - F_{(\theta^{EXT}_{BCS} \mid X_{MCS})} \right] +
  \left[ F_{(\theta^{EXT}_{BCS} \mid X_{MCS})} - F_{(\theta^{EXT}_{BCS} \mid X_{BCS})} \right]$$

% 
% <<COUNTERF, dependson="FA_SCORES_MAIN">>=
% # perform counterfactuals
% @
% 
% <<COUNTERF_PREP_PLOTS, dependson='COUNTERF'>>=
% # prepare counterfactual plots
% @
% 
% \begin{landscape}
% \begin{figure}
% <<COUNTERF_EXT, dependson='COUNTERF_PREP_PLOTS', dev='tikz', fig.height=10, fig.width=16, out.width='1\\linewidth'>>=
% cfp.ext
% @
% \caption{Counterfactual decomposition -- externalising skills}\label{fig:counterf_ext}
% {\scriptsize \emph{Notes}: \par}
% \end{figure}
% \end{landscape}
% 
% \begin{landscape}
% \begin{figure}
% <<COUNTERF_INT, dependson='COUNTERF_PREP_PLOTS', dev='tikz', fig.height=10, fig.width=16, out.width='1\\linewidth'>>=
% cfp.int
% @
% \caption{Counterfactual decomposition -- internalising skills}\label{fig:counterf_int}
% {\scriptsize \emph{Notes}: \par}
% \end{figure}
% \end{landscape}
% 
% <<COUNTERF_PREP_TAB, dependson='COUNTERF'>>=
% # prepare counterfactual tables
% @
% 
% <<COUNTERF_INEQ, dependson='COUNTERF_PREP_TABS', results='markup'>>=
% cat("Males EXT")
% print(iqdecomp(logitcf.ext.m))
% 
% cat("Males INT")
% print(iqdecomp(logitcf.int.m))
% 
% cat("Females EXT")
% print(iqdecomp(logitcf.ext.f))
% 
% cat("Females INT")
% print(iqdecomp(logitcf.int.f))
% @

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% COUNTERFACTUAL DECOMPOSITIONS

<<COUNTERF_SEQDEC_PREP>>=
# data for decomposition plots
@

<<COUNTERF_SEQDEC_PLOT>>=
# decomposition plots
@

\begin{landscape}
\begin{figure}
<<COUNTERF_SEQDEC_EXT, dependson='COUNTERF_SEQDEC_PREP', dev='tikz', fig.height=10, fig.width=16, out.width='1\\linewidth'>>=
cfp.seqdec.ext
@
\caption{Counterfactual sequential decomposition -- externalising skills}\label{fig:counterf_seq_ext}
{\scriptsize \emph{Notes}: \par}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}
<<COUNTERF_SEQDEC_INT, dependson='COUNTERF_SEQDEC_PREP', dev='tikz', fig.height=10, fig.width=16, out.width='1\\linewidth'>>=
cfp.seqdec.int
@
\caption{Counterfactual sequential decomposition -- internalising skills}\label{fig:counterf_seq_int}
{\scriptsize \emph{Notes}: \par}
\end{figure}
\end{landscape}

@@

<<BCSEFA>>=
@

<<MCSEFA>>=
@

@@

datacompl <- scores2plot[complete.cases(scores2plot[,decvars]),] # complete cases
# expand dummies
datasum <- cbind(
  datacompl[,c("EXT","INT","cohortsex")], 
  dummy.data.frame(datacompl[,c(decvars)], sep = "")
)

>>>>

\documentclass[pdftex,10pt,twoside,a4paper]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{amstext}% for resizing the symbol in math
\usepackage{amsthm}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[babel]{csquotes}
\usepackage{booktabs}
\usepackage{tocloft}
\usepackage{dsfont}
\usepackage{bbm}
\usepackage{xcolor}
\usepackage{accents}
\usepackage{setspace} % for spacing
\onehalfspacing
\usepackage{rotating}
\usepackage{fullpage}
\usepackage{pdflscape}  % for landscape page
\usepackage{lmodern}  %font
\usepackage{natbib}
\usepackage{array}
\usepackage{graphicx}
\usepackage{adjustbox}
\usepackage{tikz}
\usepackage{bm}
\usepackage{tabularx}
\usepackage{marvosym}
\usepackage{calc}
\usepackage{array}
\usepackage{verbatim}
\usepackage{multirow}
\usepackage[multiple]{footmisc}   % for multiple footnotes
\usepackage{textgreek}            % for greek letters in text
\usepackage{ragged2e}             % for table notes justification

@@

@@

@@

\textbf{REVIEW} We can derive scores for $\theta^{EXT}$ and $\theta^{INT}$ directly, and plot them against socioeconomic measures observed in the data. Our main socioeconomic indicator is a measure of weekly total family net income in the prices of the interview year.\footnote{This measure is harmonised across cohorts and derived as part of CLOSER, which is funded by the ESRC.} Using the estimated parameters for the model with invariant thresholds and loadings (but non-invariant intercepts), I compute the scores for each individual in the sample (using the Empirical Bayes modal approach).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

read_chunk(here("codes/R/1_load.R"))
read_chunk(here("codes/R/2_efa.R"))
read_chunk(here("codes/R/3_inv_main.R"))
read_chunk(here("codes/R/3_inv_aux.R"))
read_chunk(here("codes/R/3_inv_fit.R"))
read_chunk(here("codes/R/4_scores.R"))
read_chunk(here("codes/R/4_scores_aux.R"))
read_chunk(here("codes/R/5_graphs.R"))
read_chunk(here("codes/R/6_selectvars.R"))
read_chunk(here("codes/R/7_counterf.R"))
read_chunk(here("codes/R/7_counterf_graphs.R"))
read_chunk(here("codes/R/7_counterf_tabs.R"))
read_chunk(here("codes/R/7_counterf_seqdec.R"))
read_chunk(here("codes/R/8_regs_dep.R"))
read_chunk(here("codes/R/8_regs_outc.R"))
read_chunk(here("codes/R/getpars.R"))

<<PREAMBLE >>=
@

@@

<<FA_SCORES_BIND>>=
# bind scores with data
@

@@

\clearpage
\begin{figure}
<<FACDENS, dependson="FA_SCORES_BIND", dev='tikz', fig.height=10, fig.width=10, out.width='1\\linewidth'>>=
@
\caption{Density of Scored Factors}
{\scriptsize \emph{Notes}: The figure shows estimated kernel densities of the scored factors separately by cohort.  \par}
\end{figure}

\clearpage
\begin{figure}
<<RAWHIST, dependson="FA_SCORES_BIND", dev='tikz', fig.height=10, fig.width=10, out.width='1\\linewidth'>>=
@
\caption{Frequency of Raw Scores}
{\scriptsize \emph{Notes}: The figure shows frequencies of raw (sum) scores separately by cohort. Raw scores are obtained using the same 11 items in the CFA model, building a categorical index. Children for which each behaviour ``certainly applies/true" are assigned zero points, one point for ``somewhat applies/true" and two points for ``does not apply/not true". For binary items, one point is assigned for ``true". The raw scores are the sum of the points across the 11 behaviours considered, six for EXT and five for INT. \par}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[ht]
\centering
\small
\caption{Inequality measures for behavioural skills}\label{tab:ineqscores}
<<INEQ_STATS, dependson="FA_SCORES_BIND", results='asis'>>=

# income 
tb <- data.frame(
  aggregate(scores2plot$faminc10_infl, by=list(scores2plot$cohort), FUN = function(x) quantile(x, c(.1, .25, .75, .9), na.rm=T))
)
tb$d9010 <- tb$x[,4]/tb$x[,1]
tb$d7525 <- tb$x[,3]/tb$x[,2]

tb.ext <- data.frame(
  aggregate(scores2plot$EXT, by=list(scores2plot$cohortsex), FUN = function(x) quantile(x, c(.1, .25, .5, .75, .9), na.rm=T))
)
tb.ext$d9010 <- abs(tb.ext$x[,"90%"]-tb.ext$x[,"10%"])
tb.ext$d7525 <- abs(tb.ext$x[,"75%"]-tb.ext$x[,"25%"])
tb.ext$d9050 <- abs(tb.ext$x[,"90%"]-tb.ext$x[,"50%"])
tb.ext$d5010 <- abs(tb.ext$x[,"50%"]-tb.ext$x[,"10%"])
rownames(tb.ext) <- tb.ext$Group.1
tb.ext <- t(tb.ext[,-c(1,2)]) # remove aggregated quantiles and transpose
tb.ext <- tb.ext[,c("BCS.M","MCS.M","BCS.F","MCS.F")]

tb.int <- data.frame(
  aggregate(scores2plot$INT, by=list(scores2plot$cohortsex), FUN = function(x) quantile(x, c(.1, .25, .5, .75, .9), na.rm=T))
)
tb.int$d9010 <- abs(tb.int$x[,"90%"]-tb.int$x[,"10%"])
tb.int$d7525 <- abs(tb.int$x[,"75%"]-tb.int$x[,"25%"])
tb.int$d9050 <- abs(tb.int$x[,"90%"]-tb.int$x[,"50%"])
tb.int$d5010 <- abs(tb.int$x[,"50%"]-tb.int$x[,"10%"])
rownames(tb.int) <- tb.int$Group.1
tb.int <- t(tb.int[,-c(1,2)]) # remove aggregated quantiles and transpose
tb.int <- tb.int[,c("BCS.M","MCS.M","BCS.F","MCS.F")]

# assemble and print
table <- rbind(tb.ext,tb.int)
table <- cbind(data.frame(as.matrix(rep(c("90-10", "75-25", "90-50", "50-10"),2))),
               table)

print(xtable(table,
                    align="L{0cm}L{.20\\textwidth}C{.15\\textwidth}C{.15\\textwidth}C{.15\\textwidth}C{.15\\textwidth}",
                    digits = 4
                    ),
             size="\\small",
             include.rownames=F,
             include.colnames=F,
             booktabs = T,
             add.to.row = list(pos = c(list(0,4)),
                              command = c(
                              # header
            "Measure & \\multicolumn{2}{c}{\\textbf{Males}} & \\multicolumn{2}{c}{\\textbf{Females}} \\\\ \\cmidrule(lr){2-3}  \\cmidrule(lr){4-5} 
                       & BCS & MCS & BCS & MCS  \\\\
                      \\textbf{Externalising} \\\\ \\midrule ",
                      "\\textbf{Internalising} \\\\ \\midrule "
                                    )),
             sanitize.text.function = function(x){x},
             sanitize.rownames.function = function(x){""}, # to not print row numbers
             tabular.environment = "tabular", width = "\\textwidth",
             floating = FALSE
)

<<FACINEQ_MAKE>>=
# make factor inequality graphs
@

\clearpage
\begin{figure}
<<FACINEQ_SC, dependson="FA_SCORES_BIND", dev='tikz', fig.height=12, fig.width=10, out.width='1\\linewidth'>>=
# plot
fi <- fi_sc # select result
gridExtra::grid.arrange(fi$ineqlist[[1]],fi$ineqlist[[2]], 
                        fi$ineqlist[[3]],fi$ineqlist[[4]],
                        fi$histlist[[1]],fi$histlist[[2]], fi$legend,
                        ncol=2, nrow=4, heights = c(4,4,2,0.5))
@
\caption{Mean scores across social class}\label{fig:meanci_sc}
{\scriptsize \emph{Notes}: The figure plots the mean and its 95\% confidence interval across harmonised social class of the cohort member's family at 10 years of age. \par}
\end{figure}

\clearpage
\begin{figure}
<<FACINEQ_YS, dependson="FA_SCORES_BIND", dev='tikz', fig.height=12, fig.width=10, out.width='1\\linewidth'>>=
fi <- fi_ys # select result
gridExtra::grid.arrange(fi$ineqlist[[1]],fi$ineqlist[[2]], 
                        fi$ineqlist[[3]],fi$ineqlist[[4]],
                        fi$histlist[[1]],fi$histlist[[2]], fi$legend,
                        ncol=2, nrow=4, heights = c(4,4,2,0.5))
@
\caption{Mean scores across maternal schooling}\label{fig:meanci_ys}
{\scriptsize \emph{Notes}: The figure plots the mean and its 95\% confidence interval across grouped years of schooling of the cohort member's mother at 5 years of age. \par}
\end{figure}

\clearpage
\par\vfill\break % break page to change margins
\advance\hoffset by -1cm % Shift top margin
\begin{landscape}
\begin{figure}
<<FACINEQ_MECS, dependson="FA_SCORES_BIND", dev='tikz', fig.height=10, fig.width=14, out.width='1\\linewidth'>>=
fi <- fi_mecs # select result
gridExtra::grid.arrange(fi$ineqlist[[1]],fi$ineqlist[[2]], 
                        fi$ineqlist[[3]],fi$ineqlist[[4]],
                        fi$histlist[[1]],fi$histlist[[2]], fi$legend,
                        ncol=2, nrow=4, heights = c(4,4,1.5,0.3))
@
\caption{Mean scores across maternal employment and schooling}\label{fig:meanci_me}
{\scriptsize \emph{Notes}: The figure plots the mean and its 95\% confidence interval across interactions between a dummy for whether the cohort member's mother continued education past compulsory schooling age, and employment status of the cohort member's mother at 5 years of age. \par}
\end{figure}
\end{landscape}
\par\vfill\break
\advance\hoffset by 1cm % Shift top margin back to normal

\clearpage
\begin{figure}
<<FACINEQ_SM, dependson="FA_SCORES_BIND", dev='tikz', fig.height=12, fig.width=10, out.width='1\\linewidth'>>=
fi <- fi_sm # select result
gridExtra::grid.arrange(fi$ineqlist[[1]],fi$ineqlist[[2]], 
                        fi$ineqlist[[3]],fi$ineqlist[[4]],
                        fi$histlist[[1]],fi$histlist[[2]], fi$legend,
                        ncol=2, nrow=4, heights = c(4,4,2,0.5))
@
\caption{Mean scores across maternal marital status}\label{fig:meanci_sm}
{\scriptsize \emph{Notes}: The figure plots the mean and its 95\% confidence interval by marital status of the cohort member's mother at 5 years of age. \par}
\end{figure}

@@

<<SELECTVARS>>=
# define variables and formulas for regression and decomposition
@

<<REGS_DEP, dependson="SELECTVARS">>=
# run regressions explaining EXT INT
@

\begin{landscape}
\begin{table}
\caption{Quantifying the gradient -- Males}
\centering
\resizebox{24cm}{!}{
<<REGS_M, results='asis', dependson="REGRESSIONS">>=
library(stargazer)
stargazer(r.ext$BCS.M[[1]], r.ext$BCS.M[[2]], r.ext$MCS.M[[1]], r.ext$MCS.M[[2]], r.int$BCS.M[[1]], r.int$BCS.M[[2]], r.int$MCS.M[[1]], r.int$MCS.M[[2]],
          column.labels=c("BCS","BCS","MCS","MCS","BCS","BCS","MCS","MCS"),
          dep.var.caption = "", multicolumn = FALSE,
          #omit=c("Constant", paste0("region", c(2:9))),
          no.space=TRUE,
          omit.stat= c("f", "ser", "aic", "ll"),
          omit.table.layout = "n",
          single.row=TRUE,
          float=FALSE    # needs begin table outside it
)
@
}
\\[-12pt] \justify {\scriptsize \emph{Notes}: This table shows coefficients from regressions of externalising and internalising skills on potential determinants.
\par}
\end{table}
\end{landscape}

\begin{landscape}
\begin{table}
\caption{Quantifying the gradient -- Females}
\centering
\resizebox{24cm}{!}{
<<REGS_F, results='asis', dependson="REGRESSIONS">>=
library(stargazer)
stargazer(r.ext$BCS.F[[1]], r.ext$BCS.F[[2]], r.ext$MCS.F[[1]], r.ext$MCS.F[[2]], r.int$BCS.F[[1]], r.int$BCS.F[[2]], r.int$MCS.F[[1]], r.int$MCS.F[[2]],
          column.labels=c("BCS","BCS","MCS","MCS","BCS","BCS","MCS","MCS"),
          dep.var.caption = "", multicolumn = FALSE,
          #omit=c("Constant", paste0("region", c(2:9))),
          no.space=TRUE,
          omit.stat= c("f", "ser", "aic", "ll"),
          omit.table.layout = "n",
          single.row=TRUE,
          float=FALSE    # needs begin table outside it
)
@
}
\\[-12pt] \justify {\scriptsize \emph{Notes}: This table shows coefficients from regressions of externalising and internalising skills on potential determinants.
\par}
\end{table}
\end{landscape}

\par\vfill\break % break page to change margins
\advance\hoffset by -1cm % Shift top margin
\begin{landscape}
\begin{table}[!htp]
\caption{Gelbach decomposition of the EXT/INT differential by maternal education}
\centering
\input{../../../tables/med_gelb}
\\[-12pt] \justify {\scriptsize \emph{Notes}: This table shows the decomposition of the differential in externalising and internalising skills between children whose mothers are educated past the compulsory schooling age, and children whose mothers left school at the minimum school leaving age. Education of mothers is observed when the child is aged around 5 (1975 for BCS, 2005 for MCS). Col. (1) shows: the unconditional difference, i.e. the coefficient on maternal education in a regression that controls only for region of birth; the conditional difference, i.e. the same coefficient after adding explanatory variables; and the portion of the unconditional difference explained by each group of variables. Col. (2) shows the fraction of the total difference due to each group of variables. The decomposition follows Gelbach (2016), and is invariant to the order in which the explanatory variables are added to the conditional model.
\par}
\end{table}
\end{landscape}
\par\vfill\break
\advance\hoffset by 1cm % Shift top margin back to normal

<<REGS_OUTC, dependson="SELECTVARS">>=
# run regressions explaining adult outcomes
@

<<REGS_OUTC_TAB, dependson="REGS_OUTC">>=
# make tables
@

\begin{table}
\caption{Effect of behavioural skills on adolescent outcomes -- BCS}
\centering
\resizebox{\linewidth}{!}{
<<REGS_OUTC_BCS, results='asis', dependson="REGRESSIONS">>=

# collapse to matrix
bcs_outctab_mat <- do.call(rbind,lapply(bcs_outctab,matrix,ncol=11,byrow=FALSE))
table <- bcs_outctab_mat

print(xtable(table,
                    align="R{0cm}L{.25\\textwidth}C{.06\\textwidth}C{.08\\textwidth}C{.08\\textwidth}C{.08\\textwidth}C{.08\\textwidth}C{.06\\textwidth}C{.08\\textwidth}C{.08\\textwidth}C{.08\\textwidth}C{.08\\textwidth}",
                    digits = 3
                    ),
             size="\\small",
             include.rownames=F,
             include.colnames=F,
             booktabs = T,
             add.to.row = list(pos = c(list(0)),
                              command = c(
                              # header
            " & \\multicolumn{5}{c}{\\textbf{Males}} & \\multicolumn{5}{c}{\\textbf{Females}} \\\\ \\cmidrule(lr){2-6}  \\cmidrule(lr){7-11}
& Mean & \\multicolumn{4}{c}{Coefficients} & Mean & \\multicolumn{4}{c}{Coefficients} \\\\ \\cmidrule(lr){3-6}  \\cmidrule(lr){8-11}"
            )),
             sanitize.text.function = function(x){x},
             sanitize.rownames.function = function(x){""}, # to not print row numbers
             tabular.environment = "tabular", width = "\\textwidth",
             floating = FALSE
)
@
}
\\[-12pt] \justify {\scriptsize \emph{Notes}: This table shows coefficients from regressions of adolescent and adult outcomes on externalising and internalising skills, also conditioning on early childhood variables.
\par}
\end{table}

\begin{table}
\caption{Effect of behavioural skills on adolescent outcomes -- MCS}
\centering
\resizebox{\linewidth}{!}{
<<REGS_OUTC_MCS, results='asis', dependson="REGRESSIONS">>=

# collapse to matrix
mcs_outctab <- do.call(rbind,lapply(mcs_outctab,matrix,ncol=11,byrow=FALSE))
table <- mcs_outctab

print(xtable(table,
                    align="R{0cm}L{.25\\textwidth}C{.06\\textwidth}C{.08\\textwidth}C{.08\\textwidth}C{.08\\textwidth}C{.08\\textwidth}C{.06\\textwidth}C{.08\\textwidth}C{.08\\textwidth}C{.08\\textwidth}C{.08\\textwidth}",
                    digits = 3
                    ),
             size="\\small",
             include.rownames=F,
             include.colnames=F,
             booktabs = T,
             add.to.row = list(pos = c(list(0)),
                              command = c(
                              # header
            " & \\multicolumn{5}{c}{\\textbf{Males}} & \\multicolumn{5}{c}{\\textbf{Females}} \\\\ \\cmidrule(lr){2-6}  \\cmidrule(lr){7-11}
& Mean & \\multicolumn{4}{c}{Coefficients} & Mean & \\multicolumn{4}{c}{Coefficients} \\\\ \\cmidrule(lr){3-6}  \\cmidrule(lr){8-11}"
            )),
             sanitize.text.function = function(x){x},
             sanitize.rownames.function = function(x){""}, # to not print row numbers
             tabular.environment = "tabular", width = "\\textwidth",
             floating = FALSE
)
@
}
\\[-12pt] \justify {\scriptsize \emph{Notes}: This table shows coefficients from regressions of adolescent and adult outcomes on externalising and internalising skills, also conditioning on early childhood variables.
\par}
\end{table}

@@

<<BCSEFA, dependson="LOAD_DATA">>=
@

<<MCSEFA, dependson="LOAD_DATA">>=
@

@@

# expand dummies
datasum <- cbind(
  finaldata[,c("EXT","INT","cohortsex")], 
  dummy.data.frame(finaldata[,c(decvars)], sep = "")
)
